---
title: "Covid91 vaccine study Final2021"
author: "Giselle Alvarez"
date: "Spring 2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.
  
# Method
 
The first step is isolating the subpopulatoins of Moles, Females, Drug User, and LGBTQ.  
```{r}
moledata<- filter(FinalData,sex=="guy")
moledata$sex <- as.factor(moledata$sex)

```

# Moles
  
 
## Bar Chart
  

```{r}
Moles <- moledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

 # the group_by followed by summarize(count=n())

basicC <- ggplot(Moles,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Moles,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
This bar chart appears to show that approximately 340 people who were infected with COVID- 19, 57% the moles were given the drug, and about 43% that were infected got the placebo. 
  
# Nummerical Summary  

```{r}
table1 <- xtabs(~treatment + infected, data=moledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```  
  
The table shows that about 340 moles were infected with COVID 19 out of 10, 400 moles in the study. The table shows that 3.72% of moles who recvived the drug got covid, and 2.55% of the moles who got the placebo contracted covid. The table also shows, that the moles that got the vaccination were about 50% more likely to contract covid 19, this uggest that the vaccine is not effective for moles.

## Inferential Results
  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```  
  
The xhi squared table shows that the p-value of 0.0006181 is so small that it is highly unlikely that the results are by chance.  
$H_0$ the null hypothesis shows there is no differences of wheather mles can contract coivd 19 or not.
$H_0$ the alternative hyothesis is that the vaccine in fact does mae a difference in whether ot not a mole contracts covid 19. Therefore we reject the null hypothesis. 
 
The odds ratio from the fisher exact test is 1.47:1, when taking the odds ratio and computing it into a percent of effectivness by using the formula 0/(1+0)*100 you get 59.5% effecticvness. 
   
   
```{r}
table2 <- xtabs(~infected + treatment, data=FinalData)
rowPerc(table2)
colPerc(table2)
print(table2)
```
   
 The vaccine is 67% effective in Females.  
   
# Females

```{r}
Femaledata <- filter(FinalData,sex== "gal")
Femaledata$sex <- as.factor(Femaledata$sex)
levels(Femaledata$infected)
```
  
# LGBTQ
  
```{r}
LGBTQdata <- filter(FinalData,LGBTQ== "gay")
LGBTQdata$DrugUser <- as.factor(LGBTQdata$LGBTQ)
```   
  
```{r}
LGBTQ <- LGBTQdata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))

 # the group_by followed by summarize(count=n())

basicC <- ggplot(LGBTQ,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(LGBTQ,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```  
  
  
  
# Drug Users
  

  
```{r}
table1 <- xtabs(~treatment + infected, data=moledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```    
  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```    
  
The vaccine is 67% effictive in females.   
    
  
# Discussion & Conclusion
  
  
  
  
  
  
  
  
